A Minimal Demo of \texttt{glyphmaps}

### 2012-06-21 paper/demo.tex

\title{A Minimal Demo of glyphmaps}

@@

\maketitle
This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change). \texttt{glyphmaps} was written to create plots like the ones below. Each was made with \texttt{glyphmaps}. The first two plots recreate plots from \citet{wickham2011split}. The last two recreate plots from \citet{wickham2012glyph}.

To run the examples contained in this paper, you'll have to download the \texttt{glyphmaps} development package from \url{https://github.com/garrettgman/ggplyr} and install it with

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{devtools}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{load\usebox{\hlnormalsizeboxunderscore}all}\hlkeyword{(}\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: Can't find package glyphmaps
\end{verbatim}
\end{kframe}
\end{knitrout}

\section{Glyphs}
The above graphs are similar in that they are all built around glyphs. I define a glyph as a geom that visually summarises information about a subset of data. Glyphs can be quite simple, or very complex. Often glyphs contain their own internal coordinate systems. The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.

\section{glyphmaps geoms}
\texttt{glyphmaps} introduces a number of new geoms that make it easy to create glyphs. These geoms can be used in conjunction with \texttt{ggplot2} functions to create layered plots. \texttt{glyphmaps} geoms behave just like \texttt{ggplot2} geoms, but take two additional arguments, described in Section~\ref{sec:args}.

\subsection{new arguments}
\label{sec:args}
Each \texttt{glyphmaps} geom takes two new arguments, \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 and \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'reference' not found

\end{verbatim}
. \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 controls how the data will be divided into subsets to be paired with individual glyphs. \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'reference' not found

\end{verbatim}
 lets the user draw reference features with the glyphs to facilitate comparisons.

\begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 takes the name of the variable(s) to be used to divide the data into the subsets that will correspond to different glyphs. The \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 argument works just like a \begin{verbatim}

Error in eval(expr, envir, enclos) : object '.vars' not found

\end{verbatim}
 argument in a \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'ddply' not found

\end{verbatim}
 call. (In fact, the \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 argument in a \texttt{glyphmaps} geom will be passed into a \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'ddply' not found

\end{verbatim}
 call as the \begin{verbatim}

Error in eval(expr, envir, enclos) : object '.vars' not found

\end{verbatim}
 argument.)

\texttt{glyphmaps} provides methods for drawing four types of reference objects behind glyphs: boxes, vertical lines, horizontal lines, and corner points. These objects help viewers make comparisons between data inside different glyphs.

\section{Generalized glyphs}
Users are not limited to the pre-packaged glyph geoms. Any single layer plot can be turned into a glyph using the \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyph' not found

\end{verbatim}
 and \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'plyr' not found

\end{verbatim}
 functions.

The plyr function also allows users to key aesthetics to subgroups of data.

\section{Discussion - glyphs, embedded graphics, hierarchies}
Glyphs are geometric objects (i.e, geoms) designed to display information within the geom. In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system. In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.

@@

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{set.seed}\hlkeyword{(}\hlnumber{1121}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlkeyword{(}\hlsymbol{x}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{rnorm}\hlkeyword{(}\hlnumber{20}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
##  [1]  0.14496  0.43832  0.15319  1.08494  1.99954 -0.81188  0.16027  0.58589  0.36009
## [10] -0.02531  0.15088  0.11008  1.35968 -0.32699 -0.71638  1.80977  0.50840 -0.52746
## [19]  0.13272 -0.15594
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## [1] 0.3217
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{var}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## [1] 0.5715
\end{verbatim}
\end{kframe}
\end{knitrout}

The first element of \texttt{x} is 0.145. Boring boxplots
and histograms recorded by the PDF device:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }two{\ }plots{\ }side{\ }by{\ }side{\ }(option{\ }fig.show=\usebox{\hlnormalsizeboxsinglequote}hold\usebox{\hlnormalsizeboxsinglequote})}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{par}\hlkeyword{(}\hlargument{mar}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{cex.lab}{\ }\hlargument{=}{\ }\hlnumber{0.95}\hlkeyword{,}{\ }\hlargument{cex.axis}{\ }\hlargument{=}{\ }\hlnumber{0.9}\hlkeyword{,}{\ }\hlargument{mgp}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{2}\hlkeyword{,}{\ }\hlnumber{0.7}\hlkeyword{,}{\ }\hlnumber{0}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{tcl}{\ }\hlargument{=}{\ }\hlkeyword{-}\hlnumber{0.3}\hlkeyword{,}{\ }\hlargument{las}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{boxplot}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{hist}\hlkeyword{(}\hlsymbol{x}\hlkeyword{,}{\ }\hlargument{main}{\ }\hlargument{=}{\ }\hlstring{"{}"{}}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots1} \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots2}

}

\end{knitrout}

>>>>

\title{A Minimal Demo of \texttt{glyphmaps}}

@@

\maketitle
This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change). \texttt{glyphmaps} was written to create plots like the ones below, which recreate plots similar to those in \citet{wickham2011split} and \citet{wickham2012glyph}. Each plot is made with \texttt{glyphmaps}. To run the examples in this document please download the \texttt{glyphmaps} package with the code below. (Note: Install won't work until I finish documenting the namespace. But cloning and load_all() DOES work.)

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{devtools}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{install\usebox{\hlnormalsizeboxunderscore}github}\hlkeyword{(}\hlstring{"{}ggplyr"{}}\hlkeyword{,}{\ }\hlstring{"{}garrettgman"{}}\hlkeyword{,}{\ }\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Installing github repo(s) ggplyr/glyphmaps from garrettgman
## Installing ggplyr.zip from https://github.com/garrettgman/ggplyr/zipball
## Installing glyphmaps
## Error: Command failed (1)
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{glyphmaps}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: there is no package called 'glyphmaps'
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}us}{\ }\hlkeyword{+}{\ }\hlfunctioncall{grid}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}line}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{time}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{pred}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{grid.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{lon}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{x.nbin}{\ }\hlargument{=}{\ }\hlnumber{58}\hlkeyword{,}{\ }\hlargument{y.nbin}{\ }\hlargument{=}{\ }\hlnumber{25}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{avg}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{alpha}{\ }\hlargument{=}{\ }\hlnumber{0.2}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlnumber{NA}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: could not find function "ggplot"
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}nasa}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}star}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{r}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{date}\hlkeyword{,}{\ }\hlargument{x}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{major.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}long"{}}\hlkeyword{,}{\ }\hlstring{"{}lat"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: could not find function "ggplot"
\end{verbatim}
\end{kframe}
\end{knitrout}

\section{Glyphs}
The common feature of these graphs is that they are built around glyphs. Glyphs are geometric objects (i.e, geoms) designed to display information within the geom. In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system. In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems. The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.

@@

To run the examples contained below, you'll have to download the \texttt{ggplyr} development package from \url{https://github.com/garrettgman/ggplyr} and install it with

@@

The first element of \texttt{x} is \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'x' not found

\end{verbatim}
. Boring boxplots
and histograms recorded by the PDF device:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }two{\ }plots{\ }side{\ }by{\ }side{\ }(option{\ }fig.show=\usebox{\hlnormalsizeboxsinglequote}hold\usebox{\hlnormalsizeboxsinglequote})}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{par}\hlkeyword{(}\hlargument{mar}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{cex.lab}{\ }\hlargument{=}{\ }\hlnumber{0.95}\hlkeyword{,}{\ }\hlargument{cex.axis}{\ }\hlargument{=}{\ }\hlnumber{0.9}\hlkeyword{,}{\ }\hlargument{mgp}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{2}\hlkeyword{,}{\ }\hlnumber{0.7}\hlkeyword{,}{\ }\hlnumber{0}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{tcl}{\ }\hlargument{=}{\ }\hlkeyword{-}\hlnumber{0.3}\hlkeyword{,}{\ }\hlargument{las}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{boxplot}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: object 'x' not found
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{hist}\hlkeyword{(}\hlsymbol{x}\hlkeyword{,}{\ }\hlargument{main}{\ }\hlargument{=}{\ }\hlstring{"{}"{}}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: object 'x' not found
\end{verbatim}
\end{kframe}
\end{knitrout}

### 2012-06-21 paper/paper.tex

>>>>

\texttt{ply$\_$aes} wraps a layer like \texttt{grid} or \texttt{glyph} and causes the aesthetics to be calculated in a groupwise manner. Users can specify the groups to use, or \texttt{ply$\_$aes} will default to any groupings provided by glyph.by or group.by parameters, or a group aesthetic.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }normal{\ }\usebox{\hlnormalsizeboxbackslash}texttt\usebox{\hlnormalsizeboxopenbrace}ggplot2\usebox{\hlnormalsizeboxclosebrace}{\ }behavior}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{hwy}\hlkeyword{,}{\ }\hlsymbol{cty}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}{\ }\hlkeyword{+}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{facet\usebox{\hlnormalsizeboxunderscore}wrap}\hlkeyword{(}\hlkeyword{\urltilda{}}\hlsymbol{cyl}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }groupwise{\ }aesthetics}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{ply\usebox{\hlnormalsizeboxunderscore}aes}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{hwy}\hlkeyword{,}{\ }\hlsymbol{cty}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}cyl"{}}\hlkeyword{)}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{facet\usebox{\hlnormalsizeboxunderscore}wrap}\hlkeyword{(}\hlkeyword{\urltilda{}}\hlsymbol{cyl}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise11} \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise12}

\texttt{ply$\_$aes} doesn't actually create embedded plots, but is closely related to the glyphing process. By default, \texttt{ggplot2} does not calculate statistics in a groupwise manner. However, such calculations can be useful in certain situations. For example, \texttt{ply$\_$aes} is built into \texttt{glyph} and \texttt{grid} by default; embedded plots quickly become boring when aesthetics do not vary from subplot to subplot. Groupwise aesthetics can be interesting in other situations as well. For example, Figure~\ref{fig:groupwise2} demonstrates how groupwise aesthetics can be used to avoid overplotting.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }overplotted}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{data}{\ }\hlargument{=}{\ }\hlsymbol{nasa}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }using{\ }ply\usebox{\hlnormalsizeboxunderscore}aes{\ }to{\ }plot{\ }just{\ }summaries}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{ply\usebox{\hlnormalsizeboxunderscore}aes}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{ozone}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{data}{\ }\hlargument{=}{\ }\hlsymbol{nasa}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise21} \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise22}

\section{speed considerations}
Embedding plots is computationally expensive. Such plots require much of both the computational engine of R and the graphics device. Hence, it is not unreasonable to expect \texttt{glyphmaps} to be slow, which it can be. However, relative to similar visual tasks \texttt{glyphmaps} is surprisingly fast. The final code chunk compares the speed of plotting forty thousand points in an embedded plot to plotting the same points in a facetted plot. \texttt{glyphmaps} performs quite well in comparison.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }embedded{\ }plot}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{system.time}\hlkeyword{(}\hlfunctioncall{print}\hlkeyword{(}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{surftemp}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{2}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
##    user  system elapsed 
##  12.460   2.282  15.981 
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hspace*{\fill}\\
\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }facetted{\ }plot}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{system.time}\hlkeyword{(}\hlfunctioncall{print}\hlkeyword{(}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{surftemp}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{2}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{facet\usebox{\hlnormalsizeboxunderscore}grid}\hlkeyword{(}\hlsymbol{long}\hlkeyword{\urltilda{}}\hlsymbol{lat}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: argument "major.aes" is missing, with no default
## Timing stopped at: 0.012 0.007 0.019 
\end{verbatim}
\end{kframe}

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-<<speed-test}
